Why Dependent Types Matter

### 2010-10-26 papers/ydtm-rev/ydtm-rev.tex
ydtm-rev: fundam.bst (ugly but compulsory?)
james-auctex
ydtm: fix typos

More specifically, we have seen how the apparatus of dependent types can
be used to maintain the length invariant in our sorting algorithm, but
that the length can essentially be regarded as a prior static notion
which the code must \emph{respect} dynamically.  What if we wanted to
guarantee the the output of our sorting algorithm is in order? The
order is not a prior static notion---that is why we need a sorting
algorithm---the order is \emph{established} by run-time testing of the
dynamic data. Can we observe this fact statically? As things stand,
$\TC{Order}$ does not matter: we can swap around the $\DC{le}$ and
$\DC{ge}$ outputs of our $\FN{order}$ test without affecting the well
typedness of $\FN{merge}$. How can we make $\TC{Order}$ matter? By
making $\TC{Order}$ a dependent type!

@@

\textbf{First-Class Modules.}$\quad$ No programming language can succeed
without strong support for the large-scale engineering of
systems. Dependent type systems already allow us to express record
types which pack up data structures, operations over them---and also
proofs of the properties of those operations. \emph{Manifest} record
types, which specify the values of some of their fields, can be used
to express sharing between records, and between the the inputs and
outputs of record-transforming operations~\cite{pollack:records}.
Epigram's first-class notion of programming pattern allows an abstract
datatype to offer admissible notions of pattern matching which hide
the actual data representation but are guaranteed to be faithful to
it---we have Wadler's \emph{views} for free~\cite{wadler:views}.

@@

\bibliographystyle{alpha}
\bibliography{handy,obsdec,alti,local}

@@

>>>>

More specifically, we have seen how the apparatus of dependent types can
be used to maintain the length invariant in our sorting algorithm, but
that the length can essentially be regarded as a prior static notion
which the code must \emph{respect} dynamically.  What if we wanted to
guarantee that the output of our sorting algorithm is in order? The
order is not a prior static notion---that is why we need a sorting
algorithm---the order is \emph{established} by run-time testing of the
dynamic data. Can we observe this fact statically? As things stand,
$\TC{Order}$ does not matter: we can swap around the $\DC{le}$ and
$\DC{ge}$ outputs of our $\FN{order}$ test without affecting the well
typedness of $\FN{merge}$. How can we make $\TC{Order}$ matter? By
making $\TC{Order}$ a dependent type!

@@

\textbf{First-Class Modules.}$\quad$ No programming language can succeed
without strong support for the large-scale engineering of
systems. Dependent type systems already allow us to express record
types which pack up data structures, operations over them---and also
proofs of the properties of those operations. \emph{Manifest} record
types, which specify the values of some of their fields, can be used
to express sharing between records, and between the inputs and
outputs of record-transforming operations~\cite{pollack:records}.
Epigram's first-class notion of programming pattern allows an abstract
datatype to offer admissible notions of pattern matching which hide
the actual data representation but are guaranteed to be faithful to
it---we have Wadler's \emph{views} for free~\cite{wadler:views}.

@@

\bibliographystyle{fundam}
\bibliography{handy,obsdec,alti,local}

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:
